---
title: "COVID-19 condition study in the United States"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---


```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)
library(stringr)
library(tidyverse)
library(data.table)
library(knitr)
library(sjPlot)
library(ggthemes)
library(scales)
library(ggwordcloud)
library(ggpubr)
options(scipen = 200)
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = TRUE,
  fig.width = 7,
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px")
```



## Introduction 

COVID-19 is a global pandemic that affects our health and life. By exploring the COVID-19 condition, we can make take possible actions to better contain its spread and make plans for the future. In this project, we mainly study the COVID-19 condition in the United States. The data for this study coms from https://covidtracking.com/data/api and  https://covid.cdc.gov/covid-data-tracker/#demographics

How is the COVID-19 condition in the United States now? The question can be answered from the following perspectives:


## Q1: Condition Overview

Here is the summary of the latest numbers about the total tests, accumulated confirmed case numbers, accumulated deaths, accumulated recoveries. 


```{r,echo=FALSE}
#df = fread("https://api.covidtracking.com/v1/us/daily.csv")
df = fread("data/daily.csv")
overview = df[1,c("date","totalTestResults","positive","recovered","death")]
colnames(overview) = c("date","total tests", "accumulated confirmed case numbers", "accumulated recoveries", "accumulated deaths")
kable(overview)
```


## Q2: Pandemic Tendency

The tendency of COVID-19 can reflect how will this pandemic will proceed into the future.We can illustrate the pandemic tendency using three important variables: New COVID-19 Tests, New Positive Cases, New Death Cases. 

```{r, echo=FALSE,  fig.width = 9, fig.align = 'center',  out.width = "900px"}
#df = fread("https://api.covidtracking.com/v1/us/daily.csv")
df = fread("data/daily.csv")
df_melted = melt(df, id.vars="date",measure.vars=c("totalTestResultsIncrease","positiveIncrease","deathIncrease"), value.name="case")
df_melted$date = as.Date(as.character(df_melted$date),"%Y%m%d")
levels(df_melted$variable) <- c("New COVID-19 Tests", "New Positive Cases", "New Death Cases")
p <- df_melted %>%
   ggplot(mapping= aes(x=date, y=case))+geom_line(aes(color=variable))+geom_area(alpha=0.5,aes(fill=variable))+
   scale_x_date(date_breaks = "1 month", minor_breaks = "1 week", labels = date_format("%b-%d"))+
   labs(x='Date',y=NULL)+
   theme(axis.text.x = element_text(angle = 0))+
   theme(plot.title = element_text(hjust = 0.5),legend.position = "none")+
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
   theme(panel.background = element_blank(), axis.line = element_line(colour =    
   "black"))+facet_wrap(.~variable,scales = "free_y",ncol=1)+
   ggtitle("COVID-19 Pandemic Tendencies")
ggplotly(p)
```

**Insight**:

- We are doing more and more COVID-19 testing!
- More and more people getting infected with COVID-19. The curve indicates that the virus spreading is speeding up as the time goes.
- We have not reached the turning point in which the actual condition gets better. 
- The new death cases is becoming flat and not increasing like new confirmed cases.

## Q3: Community Infection Status

Different age groups may have different susceptibility towards the virus due to their different immune levels. And the virus infection may also differs with respect to races and ethnicities. 

```{r, echp=FALSE}
df_case  = fread("./data/cases_by_race_ethnicity__all_age_groups.csv")
df_death = fread("./data/deaths_by_race_ethnicity__all_age_groups.csv")
df_case$label = paste(df_case$Count, "(", df_case$Percentage, "%)", sep="")
df_death$label = paste(df_death$Count, "(", df_death$Percentage, "%)", sep="")
df_case$type  = "cases"
df_death$type = "deaths"
df = rbind(df_case,df_death)
colnames(df_case)[1] = "Race_Ethnicity"
colnames(df_death)[1] = "Race_Ethnicity"
df$Percentage = as.numeric(df$Percentage)
fig <- plot_ly()
fig <- fig %>% add_pie(data = df_case, labels = ~Race_Ethnicity, values = ~Count,
          name = "Cases", 
          domain = list(x = c(0, 0.4), y = c(0.2, 1)))
fig <- fig %>% add_pie(data = df_death, labels = ~Race_Ethnicity, values = ~Count,
          name = "Deaths", domain = list(x = c(0.6, 1), y = c(0.2, 1)))
fig <- fig %>% layout(title =  paste("Infection status of different Race_Ethnicity","Left: cases", "Right: deaths", sep=" "), showlegend = T,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = TRUE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = TRUE))

fig
```


**Insight**: 
- White Non-Hispanic has both the highest infection rate and covid-death rate, whereas Asian Non-Hispanic, American Indian, Native Hawaiian has significantly low infection rate and death rate. 


```{r,echo=FALSE}
df_case = fread("./data/cases_by_age_group.csv")
df_death = fread("./data/deaths_by_age_group.csv")
df_case$label = paste(df_case$Count, "(", df_case$Percentage, "%)", sep="")
df_death$label = paste(df_death$Count, "(", df_death$Percentage, "%)", sep="")
df_case$percent = df_case$Count*100/sum(df_case$Count)
df_death$percent = df_death$Count*100/sum(df_death$Count)
df_case$type  = "COVID-19 cases"
df_death$type = "COVID-19 deaths"
df = rbind(df_case,df_death)
colnames(df)[1] = "age"
df$age = factor(df$age)
levels(df$age) = c("0 - 4 Years","5 - 17 Years","18 - 29 Years","30 - 39 Years", "40 - 49 Years","50 - 64 Years","65 - 74 Years","75 - 84 Years","85+ Years")
p <- ggplot(df,aes(x=age, y=percent))+geom_bar(stat="identity",fill="Orange")+
    geom_smooth(aes(group=1), method = "loess", se = FALSE)+ theme(axis.text.x = element_text(angle =    45,vjust=15))+labs(x=NULL)+facet_wrap(.~type)+
   theme(panel.background = element_blank(), axis.line = element_line(colour =    
   "black"),plot.title = element_text(hjust = 0.5))+ggtitle("Infection status of different age groups")
ggplotly(p)
```

**Insight:**

- The younger people have a higher infection rate than other populations. 
- The older people tends to be impacted seriously by the virus, thus leading to the higher death rate among the population.


## Q4: Population Infection Rate
The population infection rate in different states can serve as an indicator of virus spreading level.

```{r, echo=FALSE}
#df2 = data.table::fread("https://api.covidtracking.com/v1/states/current.csv")
df2 = data.table::fread("data/states_current.csv")
df_popu = data.table::fread("./data/us_census_2018_population_estimates_states.csv")
df3 <- df2 %>%
    select(state, state, positive) %>%
    left_join(df_popu, by="state")
df3$rate = df3$positive/df3$population
df3 <- df3 %>% select(state,rate, state_name)
colnames(df3) <- c("state","rate","state.name")
df3$hover <- with(df3, paste(state.name, '<br>', "Inection rate:", rate))
p3 <- plot_geo(df3) %>%
  add_trace(
    z = ~rate, text = ~hover, span=I(0),
    locations = ~state, locationmode="USA-states") %>%
  layout(geo=list(
                  scope = 'usa',
                  projection = list(type = 'albers usa'),
                  lakecolor = toRGB('white')
                 ),
         title=paste('population infection rate(20201111) ','<br>(Hover for value)')
        )
p3
```

**Insight:**
- Different states have different population infection rate now. Some states such as are serious than others. 
- The population infection rate has reached a significant level relative to the population indensity.



## Q5: Key Information

The text-mining of the notes gathered from every state are helpful for us to know the main topics. Here, we choose to display them using word tokens and tri-grams.

```{r,echo=FALSE}
#df = fread("https://api.covidtracking.com/v1/states/info.csv")
df = fread("data/states_info.csv")
library(tidytext)
df1 <- df %>%
   unnest_ngrams(ngram, notes,n=3) %>%
   count(ngram,sort=TRUE) %>%
   top_n(20, n)

df2 <-df %>%
  unnest_tokens(word, notes) %>%
  anti_join(stop_words, by = c("word")) %>%
  count(word, sort = TRUE) %>%
  top_n(25, n)
set.seed(10)
p1 <- ggplot(
  df1,
  aes(
    label = ngram, size = n,
    color = factor(sample.int(10, nrow(df1), replace = TRUE)),
  )
) +
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 10) +
  theme_minimal()+theme(plot.title = element_text(hjust = 0.5))+ggtitle("tri-gram word cloud")
p2 <- ggplot(
  df2,
  aes(
    label = word, size = n,
    color = factor(sample.int(10, nrow(df2), replace = TRUE)),
  )
) +
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 20) +
  theme_minimal()+theme(plot.title = element_text(hjust = 0.5))+ggtitle("token-level word cloud")
p <- ggarrange(p2, p1, labels = c("A", "B"),
          ncol = 1, nrow = 2)

p
```

**Insights:**
- "PCR test" is the most important information across all states. It means that most states is mainly focusing on COVID-19 testing now.


## Summary


How is the COVID-19 condition in the United States now?

To begin with, the COVID-19 condition is not optimistic now, we can see the huge numbers of cases in the overview part.  Firstly, the infection is still continuously growing with a higher and higher growth rate. Secondly, the virus infections and impacts are different in terms of different age groups and races/ethnicities. Thirdly, The population infection rate is different across different states and has already reached a significant level of around 2% now. Finally, most states are mainly working on doing COVID-19 testing now, this is the key information we should pay attention to.





